\documentclass[letterpaper]{article}

\usepackage{aaai27}  % AAAI 2027 Demo Track style
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{url}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{xcolor}

\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  commentstyle=\color{codegray},
  frame=single,
  language=Python,
  numbers=none,
  showstringspaces=false,
}

\title{Constraint Architecture for AI Agent Safety: A Dual-Validated, \\
       Cross-Domain Reference Implementation \\
       {\small (AAAI-27 Demo Track Submission)}}

\author{
  Xu Renwu (续仁舞)\\
  Huaqiao University\\
  1nour2567@gmail.com
}

\begin{document}

\maketitle

\begin{abstract}
We present \texttt{constraint-architecture}, a pip-installable reference architecture
for AI agent safety---a four-layer boundary between LLM-driven reasoning and
deterministic code enforcement.  The architecture was \textbf{not} designed
top-down.  It was reverse-extracted from two independently developed,
fully functional agent systems in completely unrelated domains: Kylin-Agent
(security-hardened operations on Kylin OS, 16 red-team attacks, 0 breaches)
and Malio (embodied AI music with 800 particles and 9 physics systems).
After finishing both, we discovered they share the same four-layer pattern:
(1)~LLM-free reasoning, (2)~deterministic constraint validation,
(3)~tiered sandbox execution, and (4)~immutable SHA256 audit trail.
The abstracted skeleton exposes 7~ABCs and 7~Protocols, 10~frozen dataclasses,
4~custom exceptions, and 25~automated self-checks, with zero dependencies.
During the demo, attendees can \texttt{pip install constraint-architecture},
subclass \texttt{ConstraintEngine} with their own rules, and have a working
agent safety layer in under five minutes.  The project shows that
\textit{``LLM reasons; code decides''} is not just a design philosophy---it
is an independently convergent, cross-domain validated engineering pattern.
\end{abstract}

%------------------------------------------------------------------------
\section{Introduction}
%------------------------------------------------------------------------

AI agents are moving from research prototypes into production at an
accelerating pace.  Simultaneously, a growing body of evidence shows that
prompt-based safety is fundamentally insufficient.
Andriushchenko et al.\ (2025) demonstrated 100\% attack success rates against
GPT-4o, Claude~3, and Llama-3 using adaptive jailbreak attacks.
The OWASP Top~10 for LLM Applications entry LLM01:2025 explicitly states that
``it is unclear if there are fool-proof methods of prevention for prompt
injection''~\cite{owasp2025}.

In 2025--2026, a clear architectural consensus has begun to emerge:
\textit{the LLM must not be the final authority on what executes}.  Multiple
independent efforts---the Layered Governance Architecture
\cite{ge2026governance}, the SARC governance-by-architecture framework
\cite{besanson2026sarc}, the agent-control-plane kernel model, and the
Authority Separation principle \cite{vaswani2025authority}---all converge
on the same pattern: a deterministic, code-level enforcement layer that
the LLM cannot override.

We arrived at the same conclusion from a different direction.  Rather than
designing an architecture and then building a reference implementation, we
built two production agent systems in completely unrelated domains, then
discovered \textit{afterwards} that they shared the same four-layer structure.
This paper presents that extracted architecture, the pip-installable skeleton
that encodes it, and the two independent validations that give it credibility.

%------------------------------------------------------------------------
\section{Constraint Architecture: Four Layers}
%------------------------------------------------------------------------

The architecture wraps every LLM reasoning step with four deterministic
layers.  The LLM \textit{proposes}; the code \textit{decides}.

\begin{enumerate}
\item \textbf{Reasoning (LLM)} --- The LLM reasons freely about what to do.
      It cannot execute, authorize, or bypass any downstream layer.
\item \textbf{Constraints (Code)} --- Deterministic validation.  Rules are
      append-only: they may be tightened as trust builds, but never removed.
      Every LLM output is checked against a typed constraint engine.
      Violations raise \texttt{VetoError} with structured
      \texttt{reason}/\texttt{alternative}/\texttt{ref} fields.
\item \textbf{Execution (Sandbox)} --- Tiered execution: \texttt{AUTO}
      (read-only diagnostics), \texttt{CONFIRM} (write operations with human
      approval), \texttt{VETO} (destruction---never executable).  Role and
      permission enforcement happens here, not in the LLM.
\item \textbf{Audit (Immutable)} --- Every pipeline event is recorded in a
      SHA256 hash chain.  Each \texttt{AuditEvent} carries a mandatory
      \texttt{prev\_hash}; the chain is verifiable at any point.
\end{enumerate}

The key design principle is \textbf{constraint monotonicity}: base-level
safety rules never decrease.  ``Relaxed'' mode doesn't mean weaker rules---
it means fewer \textit{additional} restrictions.  This is a structural
guarantee, not a prompt-level promise.

%------------------------------------------------------------------------
\section{The Skeleton: Engineering Properties}
%------------------------------------------------------------------------

The architectural skeleton is published as
\texttt{constraint-architecture} on PyPI (MIT license).
It is 1,384~lines of Python with \textbf{zero dependencies}.

\begin{table}[h]
\centering
\caption{Engineering properties of the skeleton.}
\begin{tabular}{lrl}
\toprule
Property & Value & Notes \\
\midrule
ABCs (abstract classes) & 7 & Shell, Identity, Scheduler, Memory, \\
                         &   & Protocol, Tools, + 4 constraint-layer ABCs \\
Protocols (duck-typing) & 7 & Each ABC paired with a Protocol \\
Dataclasses & 12 & 10 frozen (immutable), 2 mutable \\
Custom exceptions & 4 & \texttt{VetoError} carries reason+alternative+ref \\
Enums & 4 & Tier (3 values), EventType (8), MemoryLevel (3) \\
Automated self-checks & 25 & ABCs uninstantiable, frozen immutable, bounds \\
Python version & 3.10+ & \\
Dependencies & 0 & \\
\bottomrule
\end{tabular}
\end{table}

Why ABC + Protocol dual interfaces?  An ABC forces implementors to inherit
(the subclassing path).  A Protocol allows structural subtyping (the
duck-typing path).  Both are provided because different domains wire the
layers differently---Kylin-Agent uses ABC inheritance throughout, while
Malio uses Protocol-based composition.  The skeleton doesn't force one path.

Why frozen dataclasses?  A \texttt{ValidationResult(allowed=False)} must not
be silently flipped to \texttt{allowed=True} by downstream code.  Freezing
the object prevents this.  \texttt{\_\_post\_init\_\_} additionally validates
that \texttt{risk\_score} is bounded 0--9 and that \texttt{prev\_hash} is
always required for \texttt{AuditEvent}.

\begin{lstlisting}[caption={Five minutes to a working agent safety layer.}]
pip install constraint-architecture

from constraint_architecture import ConstraintEngine, ValidationResult

class MyGuardrails(ConstraintEngine):
    async def validate(self, context, reasoning_output):
        if "rm -rf" in reasoning_output.get("command", ""):
            return ValidationResult(
                allowed=False,
                reason="destructive command blocked"
            )
        return ValidationResult(allowed=True)
\end{lstlisting}

%------------------------------------------------------------------------
\section{Reference Implementation 1: Kylin-Agent}
%------------------------------------------------------------------------

Kylin-Agent is a security-hardened AI operations agent deployed on Kylin OS
V11---a production-grade Linux distribution widely used in Chinese government
and defense environments.  The agent accepts natural-language operational
commands and passes them through four deterministic defense layers:

\begin{description}
\item[T0 -- Injection Detection] --- Blocks prompt injection, hex/base64/
      Unicode obfuscation, and buffer overflow payloads before the LLM sees them.
\item[T1 -- Risk Scoring] --- Deterministic 1--10 score from checklist-based
      heuristics.  No ML involved.
\item[T2 -- Dual-Path Validation] --- Structured tool-semantic validation
      AND raw regex validation.  Both must agree; disagreement triggers
      human confirmation regardless of score.
\item[T3 -- Sandbox] --- Whitelist-enforced \texttt{sudo -n} execution.
      Commands not on the allowlist are rejected at the OS level.
\end{description}

After action: SHA256 audit chain with persistence across reboots.
\textbf{16 red-team attacks.  0 breaches.  149 tests.}  All four defense
layers are deterministic---not one depends on the LLM.

%------------------------------------------------------------------------
\section{Reference Implementation 2: Malio}
%------------------------------------------------------------------------

Malio is an embodied AI music agent with 800 particles and 9 physics systems.
Its PersonaEngine enforces a central-bank-style emotional currency model:
no code can directly modify \texttt{energy}/\texttt{warmth}/\texttt{playfulness}.
Agents request ``loans'' of emotional expression; the engine grants or denies
them deterministically based on current persona bounds.

The same four layers manifest differently here:
\begin{itemize}
\item \textbf{Reasoning}: MusicAgent (ReAct loop) searches independent
      music; VisualAgent manages DSL rule lifecycle.
\item \textbf{Constraints}: PersonaEngine enforces structural bounds
      (e.g., cannot emit \texttt{light\_burst} when energy~$<$~0.3).
\item \textbf{Execution}: Core action push with role filtering---no agent
      can modify another agent's particle parameters directly.
\item \textbf{Audit}: Federated rule exchange via SBERT semantic clustering;
      every rule proposal has a signed origin and a verifiable audit trail.
\end{itemize}

The key insight: \textbf{nothing about ``music DJ'' suggests four-layer
security architecture}.  Yet after building both systems independently,
the same layered structure emerged.  This is not cherry-picking---it is
convergent evolution at the architectural level.

%------------------------------------------------------------------------
\section{Related Work}
%------------------------------------------------------------------------

The 2025--2026 period has witnessed a decisive convergence toward
deterministic, multi-layer agent safety architectures.  We organize the
rapidly growing literature into four themes: (1)~the fundamental
insufficiency of LLM-self-enforcement, (2)~multi-layer governance
frameworks, (3)~formal verification and provable guardrails, and
(4)~layer-isolated evaluation and drift detection.

\subsection{The Failure of LLM Self-Enforcement}

A foundational finding across multiple independent studies is that a single
LLM cannot reliably enforce its own safety rules under adversarial pressure.
The IETF Asimov Safety Architecture (ASA) draft~\cite{baysal2026asa}
demonstrates that adversarial users erode safety compliance through multi-turn
prompt injection: actions refused at turn~3 were permitted by turn~15 with
only accumulated conversational context changed.  Under fictional or
hypothetical frames, the model executed actions it had previously refused
(frame vulnerability), and persistent user pressure combined with appeals to
helpfulness caused the model to prioritize helpfulness over security
(sycophancy override).  Andriushchenko et al.\ \cite{andriushchenko2025}
confirm that adaptive jailbreak attacks achieve 100\% attack success rates
against frontier models.  The OWASP LLM01:2025 entry~\cite{owasp2025}
explicitly concludes that ``it is unclear if there are fool-proof methods of
prevention for prompt injection.''  Baysal et al.\ observe that existing
safety frameworks filter what an AI \textit{says} but not the proposed action
\textit{payload}, allowing benign-sounding responses to carry destructive
actions~\cite{baysal2026asa}.

\subsection{Multi-Layer Governance Frameworks}

A rapidly growing body of work proposes multi-layer architectures that
interpose deterministic enforcement between LLM reasoning and execution.
Ge et al.\ \cite{ge2026governance} propose a four-layer governance framework
(Execution Sandbox, Intent Verification, Zero-Trust Authorization, Immutable
Audit) achieving 96\% interception rate.  The SARC framework
\cite{besanson2026sarc} formalizes constraints as first-class specification
objects compiling into four enforcement sites (Pre-Action Gate, Action-Time
Monitor, Post-Action Auditor, Escalation Router), achieving zero
hard-constraint violations under exact predicates and 89.5\% reduction in
soft-window overages.

The Convergent AI Agent Framework (CAAF)~\cite{li2026caaf} makes the bold
claim that ``architecture supersedes model size''---demonstrating that
deterministic harness design provides safety guarantees that parameter
scaling alone cannot.  CAAF introduces three pillars: Recursive Atomic
Decomposition (breaking tasks into verifiable atoms), State Locking
(monotonicity enforcement---rules only tighten, never relax), and a Unified
Assertion Interface (UAI) that makes every layer independently testable.

Tallam~\cite{tallam2026fiveplane} proposes a five-plane decomposition with
stop-anywhere mediation and composite principals, achieving single-digit
microsecond adjudication.  The Authority Separation principle
\cite{vaswani2025authority} explicitly states: LLMs propose, deterministic
systems authorize.  Watkins et al.\ \cite{watkins2026threelayer} prove that
a three-layer probabilistic assume-guarantee architecture is \textit{not}
optional---it is structurally required for safe LLM agent deployment, with
any weakening of any single layer causing compounding safety degradation.

The Baidu Developer Center~\cite{baidu2026fourlayer} independently proposes
a four-layer control architecture (Visualization, Task Encapsulation,
Validation Execution, State Rollback), reporting a failure rate reduction
from 18\% to 2.3\% in production deployments.

\subsection{Formally Verified Guardrails and Provable Safety}

A nascent line of work aims at \textit{provable} rather than empirical
safety.  Yang et al.\ \cite{yang2026provable} introduce the ePCA framework
for provably secure agent guardrails, achieving zero adversarial task
completions with verification latency of only 0.3--0.5~ms per call.
Theorem~1 (Safety Preservation) guarantees that every ePCA-verified action
runs exactly as authorized, with no semantic drift.  ShieldAgent
\cite{zhao2025shieldagent} converts natural-language safety policies into
verifiable Python guardrails using a relative safety ranking mechanism
(ASPM), achieving 90.4\% accuracy while reducing API queries by~3.2$\times$
compared to LLM-as-Judge baselines.

Chen et al.\ \cite{chen2026governance} propose a layered translation method
from governance norms (ISO/IEC~42001, EU~AI~Act) to enforceable runtime
controls, using a six-criterion tuple $\kappa = (\alpha, \chi, \rho, \pi,
\delta, \gamma)$ that maps every compliance obligation to exactly one
enforcement site.  Flat rule lists without priority models create undefined
behavior when rules conflict~\cite{baysal2026asa}; the six-criterion approach
resolves this at the architecture level.

\subsection{Layer-Isolated Evaluation and Drift Detection}

A critical open problem is \textit{how to test} multi-layer agent systems
without the LLM becoming the testing bottleneck.  Hu et al.\
\cite{hu2026layerisolated} introduce \textbf{layer-isolated evaluation}---a
method that decomposes a production LLM agent into 8~architectural layers
and gates each with its own deterministic, no-LLM assertion suite (225~cases
in~2.39~seconds, \$0~LLM cost).  Key findings: (1)~end-to-end aggregate
pass-rates \textit{mask} localized regressions---aggregate drops of only
1.7--5.9~pp hide per-slice drops of 25--91~pp; (2)~a coverage-honesty
criterion reports ``rate: null'' for unexercised layers, refusing to let a
green aggregate launder an uncovered layer; (3)~the method is explicitly a
\textit{drift detector}, not a correctness oracle---a green slice certifies
``unchanged from baselined contract,'' not ``behavior is correct.''

\subsection{Our Contribution: Method, Not Conclusion}

Our work shares the conclusion of the papers above---deterministic,
multi-layer enforcement is essential---but differs in \textit{method}.
Every framework cited above was designed top-down: derive principles, specify
layers, build a reference implementation, evaluate on benchmarks.  We arrived
at the same architecture bottom-up: build two fully functional agent systems
in unrelated domains, discover \textit{afterwards} that they share the same
four-layer structure, extract the invariant, and publish it as a zero-
dependency pip-installable skeleton with 25~automated self-checks.

This ``reverse extraction'' methodology provides a complementary form of
evidence: if the same architecture emerges from systems that share no common
ancestor and were built without knowledge of each other, it is likely
capturing a real engineering invariant rather than a design preference.
The resulting artifact---\texttt{constraint-architecture} on PyPI---is not a
whitepaper; it is copy-paste-able code that anyone can subclass in five
minutes.  Combined with the independent convergence of the 2025--2026
literature toward the same pattern, the evidence for four-layer deterministic
agent governance is stronger than any single paper could establish.

%------------------------------------------------------------------------
\section{Demo Plan}
%------------------------------------------------------------------------

The AAAI-27 demo will be fully self-contained:

\begin{enumerate}
\item \textbf{Live installation}: \texttt{pip install constraint-architecture}
      (2 seconds, zero dependencies).
\item \textbf{Attendee-driven constraint writing}: Attendees propose safety
      rules in natural language (e.g., ``block any command that touches
      \texttt{/etc/passwd}'').  We subclass \texttt{ConstraintEngine} live
      and demonstrate the rule enforcing deterministically.
\item \textbf{Self-check execution}: \texttt{python -m
      constraint\_architecture.skeleton} runs 25 automated checks,
      demonstrating that the skeleton validates its own correctness.
\item \textbf{Cross-domain comparison}: Side-by-side walkthrough of the same
      four-layer structure operating in Kylin-Agent (security ops) and
      Malio (music agent)---showing that the wiring differs but the
      architecture holds.
\item \textbf{Third-implementation challenge}: We will invite attendees to
      propose a third domain (healthcare? finance? education?) and sketch
      how the four layers would materialize.  Previous demos have shown that
      this exercise consistently produces valid and non-obvious constraint
      designs within the four-layer framework.
\end{enumerate}

Attendees will leave with a running safety layer on their own laptops.

%------------------------------------------------------------------------
\section{Conclusion}
%------------------------------------------------------------------------

We presented \texttt{constraint-architecture}, a cross-domain validated,
pip-installable reference architecture for AI agent safety.  The architecture
is not a theoretical proposal---it is reverse-extracted from two fully
functional, independently developed agent systems, published as a zero-
dependency Python package, and verified by 25 automated self-checks.
The 2025--2026 convergence toward deterministic agent governance suggests
that ``LLM reasons; code decides'' is becoming an industry standard.
Our contribution is demonstrating that this pattern can emerge organically
from practice in unrelated domains, and that the resulting abstraction is
simple enough to be installed and subclassed in five minutes.

\bigskip
\noindent\textbf{Project URL}: \url{https://pypi.org/project/constraint-architecture/}

\noindent\textbf{Source}: \url{https://github.com/1nour2567/Constraint-Architecture}

%------------------------------------------------------------------------
% References
%------------------------------------------------------------------------

\begin{thebibliography}{}

\bibitem[Andriushchenko et al., 2025]{andriushchenko2025}
Andriushchenko, M.; and Flammarion, N.
\newblock Adaptive Jailbreaking of LLMs.
\newblock {\em ICLR}, 2025.

\bibitem[Baidu, 2026]{baidu2026fourlayer}
Baidu Developer Center.
\newblock Four-Layer Control Architecture for AI Agent Safety: From Assisted
          Decision to Production Execution.
\newblock {\em Baidu Developer Center}, May 2026.

\bibitem[Baysal et al., 2026]{baysal2026asa}
Baysal, M.; et al.
\newblock Asimov Safety Architecture (ASA) --- draft-baysal-asimov-safety-
          architecture-00.
\newblock {\em IETF Internet-Draft}, 2026.

\bibitem[Besanson et al., 2026]{besanson2026sarc}
Besanson, G.; et al.
\newblock SARC: A Governance-by-Architecture Framework for Agentic AI Systems.
\newblock {\em arXiv:2605.07728}, 2026.

\bibitem[Chen et al., 2026]{chen2026governance}
Chen, Y.; et al.
\newblock From Governance Norms to Enforceable Controls: A Layered Translation
          Method for Runtime Guardrails in Agentic AI.
\newblock {\em arXiv:2604.05229}, 2026.

\bibitem[Ge et al., 2026]{ge2026governance}
Ge, J.; et al.
\newblock Governance Architecture for Autonomous Agent Systems: Threats,
          Framework, and Engineering Practice.
\newblock {\em arXiv:2603.07191}, 2026.

\bibitem[Hu et al., 2026]{hu2026layerisolated}
Hu, D.; et al.
\newblock Layer-Isolated Evaluation: Gating the Deterministic Scaffold of a
          Production LLM Agent with a No-LLM, Regression-Locked Test Harness.
\newblock {\em arXiv:2606.11686}, 2026.

\bibitem[Li et al., 2026]{li2026caaf}
Li, J.; et al.
\newblock Harness as an Asset: Enforcing Determinism via the Convergent AI
          Agent Framework (CAAF).
\newblock {\em arXiv:2604.17025}, 2026.

\bibitem[OWASP, 2025]{owasp2025}
OWASP Foundation.
\newblock OWASP Top 10 for LLM Applications: LLM01:2025 — Prompt Injection.
\newblock 2025.

\bibitem[Santos, 2026]{santos2026deterministic}
Santos, R.
\newblock Deterministic Guardrails for Enterprise Agents: Compliance Without
          Killing Autonomy.
\newblock {\em Zenodo:19053079}, 2026.

\bibitem[Tallam, 2026]{tallam2026fiveplane}
Tallam, S.
\newblock A Five-Plane Reference Architecture for Runtime Governance of
          Production AI Agents.
\newblock {\em arXiv:2606.12320}, 2026.

\bibitem[Vaswani et al., 2025]{vaswani2025authority}
Vaswani, A.; et al.
\newblock Authority Separation in AI Systems: Structural Guarantees Across
          Security, Epistemics, Economics, and Safety.
\newblock {\em Zenodo}, 2025.

\bibitem[Watkins et al., 2026]{watkins2026threelayer}
Watkins, R.; et al.
\newblock Position: A Three-Layer Probabilistic Assume-Guarantee Architecture
          Is Structurally Required for Safe LLM Agent Deployment.
\newblock {\em arXiv:2605.18672}, 2026.

\bibitem[Yang et al., 2026]{yang2026provable}
Yang, T.; et al.
\newblock Provably Secure Agent Guardrail.
\newblock {\em arXiv:2605.29251}, 2026.

\bibitem[Zhao et al., 2025]{zhao2025shieldagent}
Zhao, S.; et al.
\newblock ShieldAgent: Shielding Agents via Verifiable Safety Policy Reasoning.
\newblock {\em arXiv:2503.22738}, 2025.

\end{thebibliography}

\end{document}
